\SW: II. New Gravitational Lens Candidates from the CFHTLS Discovered through Citizen Science

### 2015-05-08 doc/sw-cfhtls.tex
Up to limitations of sims section, #203

\section{Data}
\label{sec:data}
\subsection{The CFHT Legacy Survey}
\label{sec:data:cfhtls}

@@

The \cfhtls imaging is very homogeneous and has good image quality. Most
of the lensed arcs are much brighter in the $g$ band thus, deep imaging
in this band is desirable. The limiting magnitude is 25.47 for the $g$
band which goes the deepest among all of the five bands. The mean seeing
in the $g$ band is 0.78\arcsec. The zero point to convert flux to AB
magnitude for all bands is 30. These characteristics make \cfhtls ideal
to do visual inspection for finding lenses.  We use the stacked images
from the final T0007 release taken from the Terapix
website\footnote{\texttt{http://terapix.iap.fr/cplt/T0007/doc/T0007-doc.pdf}} for this
work.

@@

In the thesis dissertation of \citet{Thanjavur2009}, 9 promising and 2
low probability candidates are reported which were discovered
serendipitously. These detections were made during the visual
inspection of the \cfhtls images as part of data reduction procedures
for the Weak Lensing survey \citep{Benjamin2007}.

\citet{Sygnet2010} carried out a search for edge-on galaxy lenses in the
\cfhtls WIDE. They identified galaxies, using {\sc SExtractor}, which
had $18<i<21$ and inclination angle $<25\,^{\circ}$. After applying few more
selection criteria and visual inspection, they found about 3 promising
and a total of 18 lens candidates.

The \af \citep{More2012} was used for finding blue arc-like features in
the entire \cfhtls imaging without any pre-selection on the type of the lensing
object. This \af, {\bf an improved version of the algorithm by
\citet{Alard2006}}, measures the second order moments of the flux distribution
in pixels within small regions to estimate the direction and extent of local
elongation of features. Pixels with high values of elongation are connected to
form an arc candidate. Finally, a set of thresholds on arc properties such as
the area, length, width, curvature and surface brightness are used to select
arc-like candidates. The search was carried out in the g-band which is the most
efficient wavelength to find typical lensed features.  This sample, called
SARCS, has 55 promising and a total of 127 lens candidates which are selected
from both \cfhtls WIDE and DEEP fields. The SARCS sample consists of some
galaxy-scale candidates and mostly groups/cluster scale lens candidates. This is
because more massive systems produce arcs or lensed images with large image
separation from the lensing galaxy which are easier to detect compared to the
galaxy-scales. In the absence of a large systematically followed up
verified sample of candidates, we choose the most promising 26 systems
as our bona fide lens sample from the \cfhtls WIDE. The total
number of lens candidates in the \cfhtls WIDE alone is 108.

In \citet{Elyiv2013}, the authors visually inspected a sample of 5500 optical
counterparts of X-ray point-like sources identified in the XMM-LSS
imaging of the \cfhtls W1 field. The goal was to find instances
of lensed quasars. Their sample consists of a total of 18 candidates out
of which 3 candidates are found to be promising.

The \rf \citep{Gavazzi2014} was used for finding compact rings or arcs
around centers of isolated and massive early-type galaxies. \rf
subtracts the point spread function (PSF)-matched $i$-band images from the $g$-band images, and
looks for excess flux in the bluer $g$-band. An object detector measures
the properties of these residual blue features, and candidates which
meet the length-width ratio and tangential alignment criteria are then
visually inspected to form the final sample. \citet{Gavazzi2014}
pre-selected $\sim$638,000 targets as either photometrically-classified
early type galaxies, or objects selected to have red centers and blue
outer parts, from the T0006 CFHTLS data release catalogs. A total of
14370 galaxies were found to show detectable blue residuals, and 2524
were visually inspected, having passed the automatic feature selection
process. This led to a total of 330 lens candidates out of which 42 are
good quality (\texttt{q\_flag} = 3) and 288 are medium quality
(\texttt{q\_flag} = 2) candidates. In addition to the main well-defined
sample of \citet{Gavazzi2014}, a further 71 candidates were reported to
be detected by earlier version of the \rf or from the \cfhtls DEEP.
From the main sample of ``\rf candidates,'' the SL2S team found, during
their follow-up campaign, 33 confirmed lenses \citep{Sonnenfeld2013a,Sonnenfeld2013b}.

The work by \citet{Maturi2014} used the arc finding code of
\citet{Seidel2007} and color properties of typical arcs to optimize arc
detection. This new approach was tested on the CFHTLS-Archive-Research
Survey \citep[CARS,][]{Erben2009} which covers an area of 37~deg$^2$
only and this entire data was also visually inspected by the authors to
estimate completeness and purity of their robotic search. They found 29
candidates with the robotic search alone and 41 candidates through pure
visual inspection some of which are known from previous searches. Most
of these candidates are medium-low
probability\footnote{\texttt{http://www.ita.uni-heidelberg.de/$\sim$maturi/Public/arcs}}.

The \rf and the \af searches are the only searches that make use of a
lens finding algorithm and that has been run on the entire \cfhtls
imaging.  Thus, these are considered as our reference sample of known
lenses from robotic searches. For the purposes of transparency and to
help with the training, the volunteers participating in \sw-\cfhtls lens
search were made aware of these two known lens samples. Images
containing the systems from the \rf and the \af samples were labelled as
``known lens candidates'' in the \sw discussion forum,
TALK\footnote{\texttt{http://talk.spacewarps.org/}}, where volunteers have the
opportunity to discuss their findings with fellow volunteers and the
science team. In this paper, we refer to the sample of 330 \rf and 108
\af lens candidates as the sample of ``known lens candidates'' and the
sample of confirmed (or most promising) 33 \rf and 26 \af as the sample
of ``known lenses''. Note that the ``known lens'' sample is a subset of
the ``known lens candidates'' sample.

@@

\subsection{Image presentation in \sw}
\label{sec:data:impres}
In order to perform a blind lens search over the entire \cfhtls
WIDE, we present the volunteers with cutouts of images selected
from the survey region. We briefly describe the image presentation here
for completeness but more information can be found in \PaperOne. We use
the $g$, $r$ and $i$-band imaging from \cfhtls which are most useful for
visual identification of lenses.  We made color composite images using
the publicly-available code, \humvi\footnote{The open source  color image
composition code used in this work is available from
\texttt{http://github.com/drphilmarshall/HumVI}} following the
prescription of \citet{Lupton2004}. The color scales are chosen to
maximize the contrast between faint extended objects. These parameters
are then fixed during the production of all the tiles, in order to allow
straightforward comparison between one image and another, and for
intuition to be built up about the appearance of stars and galaxies
across the survey.

We extract contiguous cutouts of size 81.84\arcsec\ (440 pixels) including
overlapping region of 10\arcsec (54 pixels) between the neighbouring
cutouts. This  results in a catalog of $\sim$ 430 000 cutouts for the entire \cfhtls
WIDE region. The size of the individual cutout is determined by
optimising factors such as the typical angular scales of gravitational
lenses, the number of objects seen in a single cutout and the total
number of image cutouts in the survey. If a lens
candidate happens to be too close to the edge of a cutout, then the overlap
between neighbouring cutout allows a volunteer to get a clearer view
of the same candidate in at least one of the cutouts. We note that since
the images are shown randomly, a volunteer may not necessarily come
across the neighbouring cutout unless {\bf they classify} a large
number of images. This is not a problem since our user base is extremely
large and we receive multiple classifications for the same cutout.

@@

The simulated lenses are important to train citizen scientists who are
novice to the task of finding lenses but they are also crucial for
analyzing the classifications performed by the citizen scientists (more
details can be found in \PaperOne, but see \Sref{sec:swap} for a brief
summary). In this section, we describe the framework used for generating the
simulated lens sample and give details of the sample itself along with
some known limitations.

@@

With the cross-section, and the luminosity functions specified, we
calculate the expected number of sources behind a candidate lensing
galaxy using Eq.~\ref{eqn:nsrc}. We need to generate a large number of
simulated lenses (larger than the number of real galaxy lenses we expect
to find in \cfhtls) in order to have a reasonably large and diverse
training sample for thousands of \sw volunteers. Therefore, we
artificially boost the average number of sources by a factor (see
Table~\ref{tab:thresh}), which increases the occurrence of lensing. We
draw a Poisson deviate, $N_{\rm src}$ with a mean equal to the boosted
average number of sources. If $N_{\rm src}$ is greater than zero, then
this galaxy is flagged as a potential lensing galaxy.

Next, we determine properties of the background sources for every lens.
We follow similar procedures for both background galaxies and quasars.
We draw source redshifts and luminosities from the aforementioned
distributions. {\bf We note that the sources are being drawn from a much
fainter magnitude range compared to the limiting magnitude of the
\cfhtls imaging and thus, the magnification bias\footnote{ {\bf In a
flux-limited sample from a survey, sources fainter than the flux limit
end up in the sample owing to the magnification by lensing which is
known as the magnification bias. This affects the source luminosity
function and needs to be accounted for when comparing the true and
observed luminosity functions.}} is naturally taken into account.} The
source positions with respect to the lens are drawn randomly from an
area inside the caustic.  {\bf When populating the sources within the
caustics, the finite size of the background galaxies is expected to
affect the lens cross-sections to some extent. As this factor is not
critical for the purpose of our training sample, we assume the
background galaxies to be point like for simplicity.} We perform
ray-tracing for all of the $N_{\rm src}$ sources using the publicly
available code \gravlens \citep{Keeton2000} and choose sources that
satisfy our selection criteria given below. We determine fluxes of the
lensed images and the total magnification of each of the lensed source.
We draw a source randomly for which the flux of the second brightest
lensed image and the total magnification of all lensed images meet the
thresholds given in Table~\ref{tab:thresh}.

Since we want to produce realistic looking lens systems, we simulate
lenses in each of the five \cfhtls~filters. The colors of the background
galaxies are drawn randomly from the photometric CFHTLenS catalog
\citep{Hildebrandt2012,Erben2013}.  Similarly, we use a quasar catalog
from the SDSS Data Release 9 \citep{Paris2012} from which colors are
drawn to simulate quasar lenses. Next, we assumed a Gaussian
profile\footnote{This was due to an oversight. We intend to use either
an exponential or {\bf de Vaucouleurs'} profile.} for
the galaxies. The ellipticity and the position angle (PA) are drawn
randomly from within the range given in Table~\ref{tab:thresh}. The
effective radius of the galaxy is estimated from the luminosity$-$size
relation \citep{Bernardi2003} (with a redshift scaling, to account for
size evolution) given by
\be
R_{\rm eff}= 10^{0.52} \frac{L_r^{2/3}}{{(1+z_s)}^2} \, {\bf Kpc}
\ee
where $L_r=L_s/10^{10.2} {\bf L_\odot}$. On the other hand, quasars are assumed to be
point sources and the PSF, with which quasars are convolved, is assumed
to have a Gaussian profile. The full-width-at-half-maximum of PSF is
equated to that of the median seeing for every filter. The median seeing
values are taken from Table 4 of the official Terapix T0007 release
explanatory document \footnote{
\texttt{http://terapix.iap.fr/cplt/T0007/doc/T0007-doc.pdf}}.

@@

\subsubsection{Groups-scale lenses}

@@

We calculate the luminosity and velocity dispersion for the BGG and each of the
member galaxies following the same prescription as in
\Sref{sect:gallens}. To calculate the average number of sources that get
lensed by such a system, we need to calculate the lensing cross-section for each
of these potential lensing groups. The complexity in the lens models makes it
analytically intractable to calculate the size of the caustics\footnote{The lens
mass distribution determines size and shape of the caustics. Any source
located within the caustics will form multiple lensed images which is the
criteria for strong lensing. To further understand caustics, see e.g., \citet{Schneider1992}.}.
Therefore, we generate the caustics numerically using \gravlens and then
determine the area covered by the caustics. We consider only galaxies as our
background source population since group or cluster-scale quasar lenses are
expected to be extremely rare in the \cfhtls.  Following the same procedure
as described in \Sref{sect:gallens}, we calculate the number of galaxies
expected to lie behind every potential lensing group (see Eq.~\ref{eqn:nsrc}).
As before, for each background galaxy within the lens cross-section, a redshift
and an $i$-band magnitude is determined by drawing galaxies randomly from the
respective distributions (see Eqs.~\ref{eqn:ps}-\ref{eqn:ns}).

@@

We produce catalogs with lens and source properties for each of the
three types of lenses.  The catalogs typically have lens position,
redshift, magnitudes, Einstein radius, ellipticity (whenever available)
and shear (for galaxy-scale lenses only). For the background sources, we
provide the offset from the lens center, redshift, magnitudes, total
magnification, number of lensed images. Additionally, when possible,
ellipticity and effective radius of the background galaxies have also
been provided. These catalogs are available from
\texttt{https://github.com/anupreeta27/SIMCT} and simulated lens sample can be
requested from the authors.

@@

\subsection{Limitations of the simulated sample}
The simulated lens sample, although realistic, is not perfect, due to
the simplicity of the lensing models and our limited understanding of
the uncertainties in the model parameters. Comments from citizen
scientists were very helpful in order to sort some of these failures.
Here, we describe some of the cases or aspects in which the simulations
were known to have failed to look realistic. The fraction of such
simulated sample is small (roughly 5 percent)\footnote{This estimate is
based on the number of \#simfail tags from TALK, the discussion forum.}.

>>>>

\section{Data}
\label{sec:data}

\subsection{The CFHT Legacy Survey}
\label{sec:data:cfhtls}

@@

The \cfhtls imaging is very homogeneous and has good image quality. Most
of the lensed arcs are much brighter in the $g$ band, so deep imaging
in this band is desirable. The limiting magnitude is 25.47 for the $g$
band which goes the deepest among all of the five bands. The mean seeing
in the $g$ band is 0.78\arcsec. The zero point to convert flux to AB
magnitude for all bands is 30. These characteristics make \cfhtls ideal
to do visual inspection for finding lenses.  We use the stacked images
from the final T0007 release taken from the Terapix
website\footnote{\texttt{http://terapix.iap.fr/cplt/T0007/doc/T0007-doc.pdf}}
for this work.

@@

In the thesis dissertation of \citet{Thanjavur2009}, 9 promising and 2
low probability candidates were reported as having been discovered
serendipitously. These detections were made during the visual
inspection of the \cfhtls images as part of data reduction procedures
for the Weak Lensing survey \citep{Benjamin2007}.

\citet{Sygnet2010} carried out a search for edge-on disk galaxy lenses in the
\cfhtls WIDE. They identified galaxies, using {\sc SExtractor}, which
had $18<i<21$ and inclination angle $<25\,^{\circ}$. After applying few more
selection criteria and visual inspection, they found about 3 promising
and a total of 18 lens candidates.

The \af \citep{More2012} was used for finding blue arc-like features in
the entire \cfhtls imaging without any pre-selection on the type of the lensing
object. This \af, {\bf an improved version of the algorithm by
\citet{Alard2006}}, measures the second order moments of the flux distribution
in pixels within small regions to estimate the direction and extent of local
elongation of features. Pixels with high values of elongation are connected to
form an arc candidate. Finally, a set of thresholds on arc properties such as
the area, length, width, curvature and surface brightness are used to select
arc-like candidates. The search was carried out in the g-band which is the most
efficient wavelength to find typical lensed features.  This sample, called
SARCS, has 55 promising and a total of 127 lens candidates which are selected
from both \cfhtls WIDE and DEEP fields. The SARCS sample consists of some
galaxy-scale candidates and mostly groups/cluster scale lens candidates. This is
because more massive systems produce arcs or lensed images with large image
separation from the lensing galaxy which are easier to detect compared to the
galaxy-scales. In the absence of a large systematically followed up
verified sample of candidates, we choose the most promising 26 systems
as our bona fide lens sample from the \cfhtls WIDE.
(The total number of lens candidates in the \cfhtls WIDE alone is 108.)

In \citet{Elyiv2013}, the authors visually inspected a sample of 5500 optical
counterparts of X-ray point-like sources identified in the XMM-LSS
imaging of the \cfhtls W1 field. The goal was to find instances
of lensed quasars. Their sample consists of a total of 18 candidates,
of which 3 candidates were found to be promising.

\citet{Gavazzi2014} used their \rf code to find compact rings or arcs
around centers of isolated and massive early-type galaxies. \rf
subtracts the point spread function (PSF)-matched $i$-band images from the $g$-band images, and
looks for excess flux in the bluer $g$-band. An object detector measures
the properties of these residual blue features, and candidates which
meet the length-width ratio and tangential alignment criteria are then
visually inspected to form the final sample. \citet{Gavazzi2014}
pre-selected $\sim$638,000 targets as either photometrically-classified
early type galaxies, or objects selected to have red centers and blue
outer parts, from the T0006 CFHTLS data release catalogs. A total of
14370 galaxies were found to show detectable blue residuals, and 2524
were visually inspected, having passed the automatic feature selection
process. This led to a total of 330 lens candidates out of which 42 were deemed
good quality (\texttt{q\_flag} = 3) and 288  medium quality
(\texttt{q\_flag} = 2) candidates. In addition to the main well-defined
sample of \citet{Gavazzi2014}, a further 71 candidates were reported to
have been detected by earlier versions of the \rf, or from the \cfhtls DEEP.
From the main sample of ``\rf candidates,'' the SL2S team found, during
their follow-up campaign, 33 confirmed lenses \citep{Sonnenfeld2013a,Sonnenfeld2013b}.

The work by \citet{Maturi2014} used the arc finding code of
\citet{Seidel2007} and color properties of typical arcs to optimize arc
detection. This new approach was tested on the CFHTLS-Archive-Research
Survey \citep[CARS,][]{Erben2009} which covers an area of 37~deg$^2$
only, and this entire image set was also visually inspected by the authors to
estimate the completeness and purity of their robotic search. They found 29
candidates with the robotic search alone and 41 candidates through pure
visual inspection---some of which were known from previous searches. Most
of these candidates are medium-low
probability\footnote{\texttt{http://www.ita.uni-heidelberg.de/$\sim$maturi/Public/arcs}}.

The \rf and the \af searches are the only searches that make use of a
lens finding algorithm and that have been run on the entire \cfhtls
imaging dataset.  Thus, we considered these to be our reference sample of known
lenses from robotic searches. For the purposes of transparency and to
help with the training, the volunteers participating in \sw-\cfhtls lens
search were made aware of these two known lens samples. Images
containing the systems from the \rf and the \af samples were labelled as
``known lens candidates'' in the \sw discussion forum,
\Talk,\footnote{\texttt{http://talk.spacewarps.org/}} where volunteers have the
opportunity to discuss their findings with fellow volunteers and the
science team. In this paper, we refer to the sample of 330 \rf and 108
\af lens candidates as the sample of ``known lens candidates'' and the
sample of confirmed (or most promising) 33 \rf and 26 \af as the sample
of ``known lenses''. Note that the ``known lens'' sample is a subset of
the ``known lens candidates'' sample.

@@

\subsection{Image presentation in \sw}
\label{sec:data:impres}

In order to perform a blind lens search over the entire \cfhtls
WIDE, we present the volunteers with cutouts of images selected
from the survey region. We briefly describe the image presentation here
for completeness; more information can be found in \PaperOne. We use
the $g$, $r$ and $i$-band imaging from \cfhtls which are most useful for
visual identification of lenses.  We made color composite images using
the publicly-available code, \humvi\footnote{The open source  color image
composition code used in this work is available from
\texttt{http://github.com/drphilmarshall/HumVI}} following the
prescription of \citet{Lupton2004}. The color scales were chosen to
maximize the contrast between faint extended objects and bright early type galaxies.
These parameters
were then fixed during the production of all the tiles, in order to allow
straightforward comparison between one image and another, and for
intuition to be built up about the appearance of stars and galaxies
across the survey.

We extracted contiguous cutouts of size 82\arcsec\ (440 pixels), including
overlapping region of 10\arcsec (54 pixels) between the neighbouring
cutouts. This resulted in a catalog of some 430,000 cutouts for the entire \cfhtls
WIDE region. The size of the individual cutout was determined by
optimising factors such as the typical angular scales of gravitational
lenses, the number of objects seen in a single cutout and the total
number of image cutouts in the survey. If a lens
candidate happens to be too close to the edge of a cutout, then the overlap
between neighbouring cutout allows a volunteer to get a clearer view
of the same candidate in at least one of the cutouts. We note that since
the images are shown randomly, a volunteer may not necessarily come
across the neighbouring cutout unless {\bf they classify} a large
number of images. This is not a problem since our user base is extremely
large and we receive multiple classifications for the same cutout.

@@

The simulated lenses are important to train citizen scientists who may be
new to the task of finding lenses, but they are also crucial for
analyzing the classifications performed by the citizen scientists (more
details can be found in \PaperOne, but see \Sref{sec:swap} below for a brief
summary). In this section, we describe the framework used for generating the
simulated lens sample and give details of the sample itself along with
some of its known limitations.

@@

With the cross-section, and the luminosity functions specified, we
calculate the expected number of sources behind a candidate lensing
galaxy using \Eref{eqn:nsrc}. We need to generate a large number of
simulated lenses (larger than the number of real galaxy lenses we expect
to find in \cfhtls) in order to have a reasonably large and diverse
training sample for thousands of \sw volunteers. Therefore, we
artificially boost the average number of sources by a factor (see
\Tref{tab:thresh}), which increases the occurrence of lensing. We
draw a Poisson deviate, $N_{\rm src}$ with a mean equal to the boosted
average number of sources. If $N_{\rm src}$ is greater than zero, then
this galaxy is flagged as a potential lensing galaxy.

Next, we determine properties of the background sources for every lens system.
We follow similar procedures for both background galaxies and quasars.
We draw source redshifts and luminosities from the aforementioned
distributions. {\bf We note that the sources are being drawn from a much
fainter magnitude range compared to the limiting magnitude of the
\cfhtls imaging and thus, the magnification bias\footnote{ {\bf In a
flux-limited sample from a survey, sources fainter than the flux limit
end up in the sample owing to the magnification by lensing which is
known as the magnification bias. This affects the source luminosity
function and needs to be accounted for when comparing the true and
observed luminosity functions.}} is naturally taken into account.} The
source positions with respect to the lens are drawn randomly from an
area inside the caustic.  {\bf When populating the sources within the
caustics, the finite size of the background galaxies is expected to
affect the lens cross-sections to some extent. As this factor is not
critical for the purpose of our training sample, for simplicity we assume the
background galaxies to be point like when computing cross-sections.} We perform
ray-tracing for all of the $N_{\rm src}$ sources using the publically
available code \gravlens \citep{Keeton2000} and choose sources that
satisfy our selection criteria given below. We determine the fluxes of the
lensed images and the total magnification of each of the lensed sources.
We draw a source randomly for which the flux of the second brightest
lensed image and the total magnification of all lensed images meet the
thresholds given in \Tref{tab:thresh}.

Since we want to produce realistic looking lens systems, we simulate
lenses in each of the five \cfhtls~filters. The colors of the background
galaxies are drawn randomly from the photometric CFHTLenS catalog
\citep{Hildebrandt2012,Erben2013}.  Similarly, we use a quasar catalog
from the SDSS Data Release 9 \citep{Paris2012} from which colors are
drawn to simulate quasar lenses. Next, we assumed a Gaussian
profile\footnote{This was due to an oversight. We intend to use either
an exponential or {\bf de Vaucouleurs'} profile.} for
the galaxies. The ellipticity and the position angle (PA) are drawn
randomly from within the range given in \Tref{tab:thresh}. The
effective radius of the galaxy is estimated from the luminosity$-$size
relation \citep{Bernardi2003} (with a redshift scaling, to account for
size evolution) given by
\be
R_{\rm eff}= 10^{0.52} \frac{L_r^{2/3}}{{(1+z_s)}^2} \, {\bf Kpc}
\ee
where $L_r=L_s/10^{10.2} {\bf L_\odot}$. On the other hand, quasars are assumed to be
point sources and the PSF, with which quasars are convolved, is assumed
to have a Gaussian profile. The full-width-at-half-maximum of PSF is
equated to that of the median seeing for every filter. The median seeing
values are taken from Table 4 of the official Terapix T0007 release
explanatory document \footnote{
\texttt{http://terapix.iap.fr/cplt/T0007/doc/T0007-doc.pdf}}.

@@

\subsubsection{Group-scale lenses}

@@

We calculate the luminosity and velocity dispersion for the BGG and each of the
member galaxies following the same prescription as in
\Sref{sect:gallens}. To calculate the average number of sources that get
lensed by such a system, we need to calculate the lensing cross-section for each
of these potential lensing groups. The complexity in the lens models makes it
analytically intractable to calculate the size of the caustics\footnote{The lens
mass distribution determines size and shape of the caustics. Any source
located within the caustics will form multiple lensed images which is the
criteria for strong lensing. To further understand caustics, see e.g., \citet{Schneider1992}.}.
Therefore, we generate the caustics numerically using \gravlens and then
determine the area covered by the caustics. We consider only galaxies as our
background source population since group or cluster-scale quasar lenses are
expected to be extremely rare in the \cfhtls.  Following the same procedure
as described in \Sref{sect:gallens}, we calculate the number of galaxies
expected to lie behind every potential lensing group (see \Eref{eqn:nsrc}).
As before, for each background galaxy within the lens cross-section, a redshift
and an $i$-band magnitude is determined by drawing galaxies randomly from the
respective distributions (see Equations~\ref{eqn:ps}-\ref{eqn:ns}).

@@

We produce catalogs with lens and source properties for each of the
three types of lenses.  The catalogs typically have lens position,
redshift, magnitudes, Einstein radius, ellipticity (whenever available)
and shear (for galaxy-scale lenses only). For the background sources, we
provide the offset from the lens center, redshift, magnitudes, total
magnification, number of lensed images. Additionally, when possible,
ellipticity and effective radius of the background galaxies have also
been provided. These catalogs are available from
\texttt{https://github.com/anupreeta27/SIMCT} and the
simulated lens image sample is available from the authors on request.

@@

\subsection{Limitations of the simulated lens sample}

The simulated lens sample, although realistic, is not perfect, due to
the simplicity of the lensing models and our limited understanding of
the uncertainties in the model parameters. Comments from citizen
scientists were very helpful in order to sort some of these failures.
Here, we describe some of the cases or aspects in which the simulations
were known to have failed to look realistic. The fraction of such
simulated sample is small (roughly 5 percent)\footnote{This estimate is
based on the number of \#simfail tags from TALK, the discussion forum.}.
